
\section{Rik and Andy go to Plop (an abject
Failure)}

(Rik's were written in van on way back across Europe, typed up by Jarv.
Proofread by Rik (hardcopy) + corrections added by Jarv.)

The mission was simple -- venture into Sys Mig, traverse the gaping
holes on \emph{Level 2} to reach the ratty old rope for Faulty Towers
and push into \emph{NCB}. Once there we hoped to bottom the fabled
`Plop', a big pitch just off \emph{NCB}, rumoured to be over 50 m,
strongly draughting and utterly jinxed!

Armed only with a vague description from a rather drunken Tetley the
night before, we set off for \emph{M16} during a brief lull in the
raging storm. Once down in the cave we quickly hopped up to \emph{NCB},
reviewing the excellent tourist trip across the big stuff in the system.

When we got to \emph{NCB}, we stumbled as Tetley had not mentioned the
lairy traverse over a \textasciitilde{}20 m drop on tatty 13-year-old 9
mm. We concluded that Plop was the pitch immediately below the rope from
Torn T. This was the error which cost us the pitch. The bolts continued
down the pitch and I had a sinking feeling as I reached the bottom of
this pointless lead.

After this we inspected the rest of \emph{NCB} going East, crossing the
bad traverse with some care. Andy and I took it in turns to examine the
side passages and one of the ones I inspected was, as in the legend of
'95, a very windy squeeze, which could be depth tested by throwing rocks
with some difficulty. I was sure this was it, but Andy had an earlier
memory which led us to think that it might be Godzilla. Stones took
around three or four seconds to drop!

We left with five hours to spare before callout, on the very cautious
side, and left the tackle bags at \emph{NCB}. Tomorrow we're going back,
and this time Plop must be conquered!

\name{Richard Venn}
