\textbf{Camp \emph{X-Ray} (-550 m): Inside the Hollow Mountain}


\section{Underground Logbook}

\sidenote{Found in 'AggregateofMig2007-2010.doc' : Jarv Typed up?}

\textbf{Vodla Sled 2010} \textbf{The} \textbf{Return to Camp
\emph{X-Ray}}

{[}Logbook cover{]}

\textbf{22-7} \textbf{Nick}

After about 6 hours of caving finally made it down. Met Gergely and
James on the way down as they were leaving the cave. Last bolt before
camp is horrible - needs rebolting/rerigging - 15cm lower would be
awesome. Built a tent at the camp. Required stone movement. NEED WEED!
Should have thought of it before. Listening to Massive Attack and
getting raptured. Oh yeah! Kate setting up sleeping space, Jarv went to
get more water. Camp is getting established. Looking forward to Worm's
World Party. Mike - cooking. Weed is really a missing resource. So far
so good. About 5 metres from camp is a hole with water in it, able to
hear, quickly got established as peeing corner, hope it's not a
lead\ldots{}

\textbf{23/7/10 Jarv}

Nice snooze - super warm. Nicola snored like a trooper - just a few
minutes into the classic Black Adder session. Broken sleep - probably as
Nicola got up for 2x piss. Awoken at 10:30am by the beasts crawling up
towards our pits. Tetley and Myles rustled up some hot-choc, then
wandered off down the continuing passage.

\textbf{23.7.10 2:10pm Myles}

Entered Gardener's World at 6:20 am. Made our way through, rerigged
\emph{Zimmer} on the way down. Arrived at camp at 10:30am and awakened
Jv, Mike, Kate and Niko. Wandered down Friendship Gallery for an hour or
two. Found nice lead, will investigate later. Sleep now.

\textbf{23-7 2:20pm Tetley}

It's good to be in a sleeping bag at Camp \emph{X-Ray} - seven years
after the last camp here. It's very comfy - I like the tent - some
things don't change though, Blackadder on the sound system, smash and
tuna etc. Hopefully we'll get some good pushing in tomorrow!

\textbf{23-7 6:20pm Tetley}

James and Dan arrive for a quick visit before heading off to push the
Muddy Window.

\textbf{8:20pm} Andy and Gergely arrive - I ignore them!

\textbf{23-7 10:30pm Myles}

Fucking body won't fall asleep! Must have only had a couple of hours at
most since Dan arrived. Gergely and Andy turned up at 8ish and now they
have checked out Leopard a while. Tetley's bodily functions are out of
control! May bring some corks down for his digestive tract next time.
Anyway, now for some food and tea and hopefully can stay awake till
bedtime at noon!

\textbf{23-7 11pm Tetley}

Myles and I share breakfast/dinner with Gergely and Andy. Fine food!

\textbf{24-7 12:20 James}

Breakfast with Tet and Myles. Dan and I will visit the lead we killed
off yesterday (Muddy Window) and survey it, then to Red Cow.

\textbf{24-7 1:30pm Myles}

Back in Camp for 2\textsuperscript{nd} night. Pushed Tolminska today,
good lead. Surveyed till 8am. Some nice pitches, covered in mud.
Listening to strange foreign music.

\textbf{24-7 2:05pm Tetley}

Great push down Korita today - 8 bolts, surveying etc. It's GOING,
GOING, GOING\ldots{}. Go THERE! (But try and avoid rigging future
pitches in or near the water\ldots{}..) Andy and Gergely have left to
push Leopard - James and Dan to survey Muddy Window and then go for a
jolly below Big Rick. It's been a great day - thanks Myles. Time for a
decent sleep.

\textbf{24/7/2010 23:20 Tetley}

James and Dan return on a high! 9hrs good kip in bed - I feel good!
Forgot to say I had a shit yesterday\ldots{}.

\textbf{24.7.10 James}

James and Dan went to Smash. Cool trip! Left \textasciitilde 80 m 9 mm
by Red Cow.

\textbf{25-7} \textbf{Sunday! 2.15am} \textbf{Tetley}

James and Dan in sleeping bags. Myles and I contemplating putting our
furries on. Still no sign of Andy and Gergely - hope they're OK.

\textbf{25 July '10 3:15am Andy}

AJ and GA turned one crappy lead (Leopard) into lots of great leads.

\textbf{3:40} \textbf{am} \textbf{Tetley}

Excitement mounts as Andy and Gergely return with news of BIG
DISCOVERIES! Now the day train is close to crashing (in their sleeping
bags). Myles and I are about to set off for sunrise and beer. Can't wait
to return!

{[}Andy's sketch survey, \emph{Zimmer} to end of \emph{Prince Consort
Road}{]}

\textbf{25.7.10 T- 5hrs to surface Myles}

First camp draws to an end. Been extremely fun if a little cold
sometimes. Will definitely return to explore my little rift, Tolminski
lead and Andy and Gergely's big leads! I will be back!

\textbf{25.7.2010 Gergely}

Route after traverse needs to be marked! Crystals on ceiling. Take care
of them. p.s. See you in Hotel Tolminka!

{[}Gergely's sketch topo of Leopard{]}

\textbf{Martin}

MM and WF. Gas running low - not able to get full blast, off to Leopard.

\textbf{26-7 Kate}

Arrived at camp again. Hellish trip.

1. Pulled Quad muscle. OW.

2. Got big chunk of hair jammed in descended. Started slashing hair
until Jarv rescued and undid descender.

3. Fucking toothbrush pierced magic drybag and the bristles soaked in
mud.

4. Nearly died slipping down Pink Series pitch. Fell a bit. Shook me up.
David Bowie made things better. Hopefully lots of exciting exploration
tomorrow.

{[}Kate's drawing of toothbrush stuck in mud{]}

{[}Martin's sketch topo of \emph{Cheetah}{]}

\textbf{Martin}

After a day at Leopard I have nothing to show for it. I have removed the
backup round the boulder on the floor and made a rope sling fit round a
rock bridge, which is going nowhere. After the second bolt there seems
little solid rock for a freehang and the walls are full of sharp stal.
The bolt William put in failed as the rock cracked and I was not
prepared to slice the rope by swinging back and forth looking for a
place especially as I was below the rub points.

\textbf{26-7-10 10am} \textbf{Jarv}

Awoken by Martin and William returning from their post-lunch look at
Korita (8am). Off with Mike and Kate to have our own attempt at the
apparent death trap on the far side of Leopard. Could really do with
some rawls - have drill but no battery, Arg! Will try our best and pop
back for lunch and turds.

\textbf{26-7-10 Jarv}

Leopard rerigging: Tough! 3 bolts in - 2 failed, one blew open rock, other cracked and demoted to 'deviation' class.

{[}Jarv's sketch topo of Leopard{]}

\textbf{Martin McGowan}: Rename it \emph{Cheetah} as you're always
cheating death?

So it's not finished but it is \textasciitilde 50\% safe. Needs lots of
rawl bolts to fix. Anyhow, it was good enough to get Kate and Mike down
so we went off to explore\ldots{}

{[}Jarv's sketch, A tourist guide to \emph{Wonderland}{]}

{[}Jarv's sketch showing Prince Consort Rd and \emph{Serpentine}{]}

{[}Jarv's sketch showing \emph{Serpentine}{]}

{[}Jarv's sketch showing \emph{Serpentine}{]}

\emph{Serpentine} is pitch down from Prince Consort Rd.5. Exit chamber
over/under big boulder and enjoy \textasciitilde 50 m lovely meander
with wet cascades (care!) Ending in dubious freeclimb to pitch head
above ledge and continuation of big pitch. We threw the only stone.
Rattle \textasciitilde 8s. Est 40-60+m

\textbf{Kate}

Smoked Mackerel fucking delicious. Just came back from Leopard and
\emph{Wonderland}. Pretty exciting stuff. Still a bit cranky, probs
hormonal or just reluctant to be down here after just 1 day on surface.
Being the first one at the bottom of newly rigged pitch was immense,
definitely worth any grimness suffered or to be suffered.
\emph{Wonderland} is magical, like a child's playground but with big
holes and sharp jaggedy rocks. After new pitch gorgeous little streamway
- quite small but not uncomfy. \emph{Serpentine}. Stopped at promising
pitch and did some surveying. This caving lark is pretty good. Good
times dancing with Mike and making mud covering on rock with `K' spelt
with stones. Getting out tomorrow, wonder how long it will be this time
till I'm back\ldots{} Martin just shat behind the tent and it smells
horribly awful.

\textbf{27-7 10:50am} \textbf{Tetley}

Back again (with Myles) - 55 hours and only 1 sleep since leaving.
Smooth exit out on the 25\textsuperscript{th}. Then Tolmin - 2 steaks!
Rush of enthusiasm as I walked up from Ravne - decided to come straight
down. Stayed up and packed after surface cavers went to sleep. Smooth
3hrs down. We then went to Leopard extensions. Now bed!

\textbf{27-7} \textbf{Sometime around 11am} \textbf{Myles}

Back in Camp \emph{X-Ray} after 51hrs on top. Couldn't resist Tetley's
proposal of 2\textsuperscript{nd} camp! Feels good to be back. I'm
gonna' fuck this cave up.

\textbf{27-7 11am Jarv}

Jarv, Mike, Kate To the surface once again! Lovely stay @ Camp
\emph{X-Ray}, made much more entertaining by Martin's situation comedy.
Leads are fantastic, Leopard is passable (drill and rawls to sort
fully). Somebody go there!

\textbf{27-7 10:30pm Tetley}

Great night's kip! No cavers here to wake us up\ldots{} Forgot to
mention that the camp stank of piss when we arrived\ldots{} Smell has
gone now\ldots{} maybe we've just got used to it. Installed a piss Daren
Drum (Heavy Metal). Location behind tent.

\textbf{27-7 Myles}

Where are all the cavers?

\textbf{28-7 00:10 Tetley}

We're now leaving to Korita. Tet and Myles.

\textbf{28-7 12:55 Tetley}

Back from a storming caving trip! We met Nicolas and Gergely at
\emph{Zimmer} - they slept while we pushed\ldots{} Another 7 bolts added
to BLACK KNIGHT pitch at end of Korita. Once down the stream flowed into
a duck. ``There will be a bypass'' I declared\ldots{}. and there was!
The stream passage continued -- we stayed high. Eventually
\textasciitilde 15 m drop to stream. But Myles found a dry passage --
strongly draughting -- that goes to a \textasciitilde 10 m pitch. This
dry route we called SIDEWINDER. 130 m in survey book. We left
\textasciitilde 25 m of 9 mm rope at bottom of Black Knight -- nothing
else left there. Great trip -- we returned to wake Gergely and Nick.
James and Jan are somewhere down Big Rock -- hope they're OK and back
soon\ldots{}

\textbf{28-7 1pm Myles}

Returned from this storming push with Tetley. We now have two pretty
beastly leads. Was a bit tired during final rigging of Black Knight but
perked up as soon as we found new passage. Now in underground camp.
Thought we'd killed it all when we found the duck but Tetley saved us
with his bypass seeking skills. Had to heat up Nick's willies over the
gas stove until they were flexible enough to fit on his clodhoppers. Him
and Gergely are off to push now.

\textbf{28-7 3:15pm} \textbf{Tetley}

James and Jan said they'd be back from a Red Cow push at 11am. Got woken
at 2pm. We (Gergely, Myles, Nicolas, Tet) got rescue kit
together\ldots{} grrr\ldots{} I got out of my warm pit -- put caving
gear on. Gergely and I went to find them. Glad to hear ``Yes we are OK''
when I shouted down\emph{Big Rock Candy Mountain}. It's good to hear the
leads still going -- good too to see Jan again at u/g camp! So time to
crawl into my pit (again!) for some sleep. I really need it (the wee
dram should help).

\textbf{29.7.2010 James}

Just woke up after an epic in Republika. Our sleep addled brains thought
we'd have enough time for a deep push\ldots{} needless to say we just
ended up sleeping all the way: on top of survey stations\ldots{}. On the
good side the drill does make rigging easier. Also the passage still
goes to the max, a 35 m pitch, 30+ m of stream. Only stopped due to lack
of ropes. Stream keeps going. Also both pitches below the chamber have
dry alcoves. PUSH!

\textbf{29-7 2:05am} \textbf{Tetley}

A broken night's sleep\ldots{} James snored like a pneumatic drill --
he's becoming my nemesis! (Only joking!) Woken by Dan + Iztok + Gergely
+ Nicolas. 8 people here it's all very jolly. Bob Dylan on the sound
system, cooking, spliff smoking (by the new comers).

Tetley's top tips

1 Pish before you shit

2 Pick your nose before sleeping

3 Use a lump of carbide and a plastic bag to dry survey instruments.

t.b.c!

\textbf{TH 29-7: 2.45 am Dan}

Finally into our sleeping bags. Izi and I came down 17:30ish. Got to
F.G. \textasciitilde 8pm. We checked out a few leads but no survey gear.
Bumped into other day (ish) train. back to camp \textasciitilde 1:30.
delicious food. Now sleep!!!

\textbf{29.7 3am} \textbf{Gergely}

Good day of pushing with Nicolas; Rolling Stones chamber and then Hidden
Surprise passage. Did some bolting, now only a couple of
life-threatening rubpoints left. Tomorrow is the day for the pitch at
the end of Hidden Surprise. Camp is very home-like feeling, nice music,
carbide light, and 8 cavers -- an underground paradise! Except for the
piss smell. Well, but everything has its downsides\ldots{} This is going
to be my longest caving trip. Whoo! Now a good sleep. Funny day train
steadily drifts towards night train. It is said that our natural day is
20-27 hours long\ldots{}

\textbf{29-7 13:45 Dan}

About to get changed. Procured survey gear (thanks Tetley) and drew
survey pages Off to push end of W.L. (If night train didn't get there
first -- still not back).

Dan's top tip

Roll up the edges of the composting bag before taking a shit.

\textbf{2/46 29/7/10 Thurs Jan}

Woke up at u/g camp to the sound of water (lots of water), went to check
out \emph{Zimmer}, and yes, there was lots of water. James hopes his
little tent is OK.

\textbf{30/7/10 Fri James}

Yesterday I took Jan down horror show: Esoterica\ldots{}

\textbf{03.54 30/7.10 Dan}

Amazing pushing trip. Found the palace of King Minos. back
\textasciitilde 1:30. Others back just in time. Good day

\textbf{Gergely}

Made traverse at end of Hidden Surprise. Other end (cont. passage) is a
strange squeeze. Probably we were the first and last persons
there\ldots{} Good company! Nice discoveries! Good cave.

{[}Gergely's sketch showing Mudstone squeeze, Hidden Surprise, Rolling
Stones{]}

We did not go down to pitch. Far side is good rock; start from 2 bolts
on traverse. End of squeeze is still going, slight draught. Bottom of
pitch is better! Waits for you!

\textbf{Fri 10:40pm 30/7/10 Jan}

So, my amazing camping trip comes to an end, its been an exciting time,
it started with a 14hr trip to Red Cow -- pushing 90 m leaving a nice
streamway still going at one of the deepest parts of Gardeners World,
shared camp with Tet + Myles (good to be back in camp with Tetley-san).
Day 2 (Thurs). I bolted a traverse over to a window in Albert Hall
(above \emph{Serpentine}), didn't go. Looked at hole beneath PSS12
(Prince Consort) pushed awkward rift to a couple of pitches ESOTERICA,
left it with two pitches undescended and looking good still. We looked
at muddy climb in Albert Hall (which turned out to be KING MINOS PALACE)
but left it for Dan + Izi, as we heard they were going to push it,
all-in-all a good introduction to the LEOPARD/PRINCE CONSORT RD series!
Back at camp we had the place to ourselves, but were expecting Jarv and
Jana., about 10pm I woke up and heard lots of water (thought it was
flooding up Friendship gallery!) and it stayed like that for the next
24hrs. Day 3 (Fri), started at 2am with Dan, Ezi, Gergerly and Niko
returning, and tales of storming passage and the maze of KING MINOS
PASSAGE! Which sounded incredible, we decided to go and see what it was
like and check some leads left unpushed, we tried to understand Dan's
detailed survey diagrams but it was all a bit complicated and
disconnected! This what I drew for Gergely after we came back\ldots{}

{[}Jan's sketch showing Palace of King Minos and the Queen's Bed
Chamber{]}

An amazing bit of cave! James looked down leads off MINOTUR RIFT and a
muddy hole with the sound of stream. In the Queen's Bed Chamber the
draught must go up I think\ldots{}? So after 3 days I'm feeling pretty
broken, but u/g camp is amazing. James and I are heading out soon, but
I've lost track of time, at some point in the last 72 hours we've
changed trains, but I don't remember the station, only the destination.
Thanks for an awesome trip James!

{[}Jan's sketches of \emph{Insomnia} series and Esoterica{]}

U/G Bivi tip \#4

Put gas canister in sleeping bag to help Butane burn. (but take it out
of your bag before trying to use it)

\textbf{Sat (morning?) 31.7.2010 Nick}

So, my second trip to UG camp turned out to be quite long and
interesting. Gergely introduced me to the extreme caving = exploration
of unknown. Tried to remember all the stuff he had told me, but not so
easy. During the time we've been down here we've pushed
\emph{Wonderland} to chamber now known as Rolling Stones (watch where
you step) . Continues with Hidden Surprise (crawl/walking passage with
chamber in the middle). Then Gergely put up a traverse (I wasn't much
help with putting in the bolts) over an unpushed (\textasciitilde 40 m)
pitch (LEAD!), once we had the traverse, it was kinda obligatory to
check out the squeeze on the other side (now known as Mudstone squeeze).
Had to take harness off and several times hammer action was needed.
After about 40-50 m, the squeeze gets even squeezier, we were quite
tired, not much draught so we turned around. Surveying was pain. Went to
sleep (3\textsuperscript{rd} night in Camp \emph{X-Ray}). Plan was to
get out but kinda didn't happen as there was too much water in
\emph{Zimmer} and presumably in higher pitches as well, so we decided to
miss callout and stay one more night. No pushing, just tourist caving,
that was the plan of the day. Went to check out passages discovered day
ago by Dan and Izi. Very Beautiful! Looks like it snowed there but
upside down, the crystal formations are crazy, definitely have to come
down with camera for photo trip. Jan and James stayed down here as well,
and day and night train kind of blended together. Now after 4 nights in
the underground we are finally going up. Thanks to Gergely the diet and
food down here was excellent, everything tastes better in the
underground!

\textbf{31-7-2010 Gergely}

Our fourth night is over. We eat up the remains of our food, drink the
last teas and hot choc. There is no way back now; we must go up today!
The water seems to be a bit less than yesterday. Probably I will wear
all thermals; will be good when waiting for Nicolas at the top of
pitches. I wonder whether there was ever a 5-day underground trip on
Mig\ldots{}

{[}Gergely's sketch showing unpushed leads in \emph{Wonderland}{]}

\textbf{31-7-2010 Gergely}

Chamber at the end of Prince Consort Rd has been renamed Albert Hall.

\textbf{1-8-10 Jarv}

Back to camp! Jarv, Jana, Martin and Shed. Arrived after the flood, had
a spot of tea to start things off -- then photo gear to Leopard. Amazing
crystals and formations, unbelievable they are there really -- only
managed to photo a fraction of them. Back to camp for
\textasciitilde 1am, noodly-mushroom-smashy goodness. Camp was a bit
trashed -- dirty crockery, dirty water, spare turd loose in its plastic
bag on the walk to camp. The previous tenants clearly also had some
obsession with burning things -- piles of ash everywhere, including
cardboard from the mess tins.

\textbf{1-8-10 Jana}

After 28h raining on top, our trip to u/g camp was postponed for 2 days.
Seeing all this new stuff, makes me feel I am in an old, exotic cave,
somewhere on a moon Billions of crystals everywhere! Now off to push
bottom of Korita.

\textbf{12:30 1-8-10 Jarv}

Martin and Shed have departed after two rounds of tea then hot choc.
Jana and I are absorbing noodles and listening to Fairport Convention.
Shared a single shit bag with her -- the relationship that shits
together, stays together! Off to Push/Photo Korita/Black Knight. Back by
23:58.

p.s. Fixed external batteries for speaker and reflexed spine of this
fine tome -- both are thus delicate!

\textbf{1-8 23:15 Tetley}

Back to Camp \emph{X-Ray}! Karin and I had a smooth 3hr journey down --
meeting Shed and Martin near the entrance and Jarv and Jana at
\emph{Zimmer}. Good food cooked by Jarv. Looking forward to a good
night's sleep then pushing tomorrow.

{[}Tetley's poor cartoon, Camp \emph{X-Ray} by Day/Night{]}

{[}Karin's excellent drawing ``Someone somewhere in the cave''{]}

\textbf{2-8-10 Jarv}

Jarv and Jana. Pushed Sidewinder - to connection with envy! Great push
-- arrived in tiny crack in ceiling (perfectly straight 10 m), with
phreatic body tube that lead to challenging (?) climb of
\textasciitilde C3-4 m down to envy pitch. ***We derigged the Sidewinder
+ a bit pitch*** you'll want \textasciitilde 10 m tat with
hanger/million for backup and massive sling for natural hang.

{[}Jarv's sketch showing extended elevation from Sindwinder.8 to envy
pitch via the Crack in Time{]}

We photo'd to the best of our ability: A bit of Sindwinder, the duck
from the far side, Black Knight - bottom and middle -- Korita middle.

Also day before: Albert hall, Muddy/NCB bit, crystals (Martin). With
Martin/Shed -- Minotaur Rift, pretties on way to Queen's Bed Chamber,
Stals on \emph{Cheetah}, \emph{Cheetah} itself.

Jana's damaged her wrist, so its to the surface we must go.

Jarv's top leads: Wet pitch at end of Black Knight -- traverse along
rift to hang? \emph{Serpentine} -- big pitch, rock looks good -- big
ledge @ 20-30 m depth.

\textbf{2-8 10:20am Tetley}

Good night's sleep in the old `buff bag' -- nice to think I was sleeping
in the same bag that we used at the Hotel in 1998! Camp \emph{X-Ray} is
becoming more like home.

\textbf{2.8} \textbf{Izi}

Štartal smo ob 4\textsuperscript{30} z bivija. V kamp smo pršli mala pr
zgoda tko de smo šli pogledat kristale. Ka smo pršli v kamp so bli
Tetley, Karin, Jana in Jarv že zbujeni, tko da smo si nardil pašta, mala
pobluzli in šli spat. Zvečer gremo phat še mala.

\textbf{2-8-10 23:58 Tetley}

Karin and I had a great day looking at all the new finds beyond Leopard.
I put 3 more bolts on \emph{Cheetah} pitch plus deviation. We looked in
vain for more easy horizontal passage. Hopefully Tjaša, Erik, Izzy and
Mawr have more luck. Time for sleep now -- James and Dave complete the 4
bed camp.

\textbf{3.8.10 James}

Went to Ballamory with Dave, the plan was to free a boulder at the
bottom of Ballamory pitch. Unfortunately more rope than we had was
necessary.

{[}James' sketch of Ballamory pitch{]}

I think we need \textasciitilde 50 m of rope. Also looks like there is a
lead before the weird doubling back bit. On the way back we stopped in
\emph{the Fridge} and rescued some carry mats. caving near Cactus
Junction is beautiful! Thanks Dave.

\textbf{3-8-10 10:15am} \textbf{Tetley}

A good night's kip. Erik, Mawr, Tjaša and Izzy have returned with news
of more discoveries. Found Rik on the stereo -- good to hear him again.
Hvala Karin! (for a very pleasant few days!) back upstairs so hopefully
the sun will be shining -- my vitamin D levels are getting low!

\textbf{3.8.10 Tjaša}

Dans smo šli obstickat minotaura in odspodi odkril še povodnega moža. Za
urško je pa štrika zmankalo Malo se bomo naspal in šli vn. Lepo smo se
mel. Hvala za dobr kamp Maur, Izi, Erik, Tjaša.

\textbf{4-8-10 Nick}

Came to Camp \emph{X-Ray} at around 2pm, faffed around a bit and went to
show Thara the newly discovered passages. Took camera with us and took
pics of the nice crystals, bumbled our way through \emph{Prince Consort
Road} to Albert hall, where we decided to turn round. Nice sightseeing.
Pushing tomorrow. Hopefully. Since no one else came on the day train, we
got the whole tent to ourselves. Noodles and fish for dinner, Sadni Čaj
for Good Night. Camp \emph{X-Ray} is as lovely as ever.

\textbf{5-8-2010 10pm Myles}

Arrived yesterday/today at 6am, went and pushed Korita series with Jarv.
Trying to kill it but is keeps going! Jarv was a tad broken before we
went to sleep. I will be tonight! Nick and Thara just returned, Nick has
a 100 m\textsuperscript{2} hole in his oversuit. Left Tetley in a
drunken state on the surface.

\textbf{5-8-10 Thara}

Left for Kamikazi pushing front. Rigged down the 10 m pitch. At the
bottom, there is a flowing passage which will be big enough for humans
in 1 million years. About 1 m above the bottom there is a window which
turns out to be a lot of crawling/squeezes. There are two small chambers
along the way for us to stand. At the second chamber we climber down
into active water passage. Upstream is dead but downstream is blocked by
a squeeze worthy of Jana. We left unsurveyed as we forgot the tape. Next
we checked the window at the level of deviation. Same thing. It kept
going and the gap on Nico's oversuit widen to cover both buttocks. On
the way out of the window, I heard a sudden change in water flow. It
must be raining outside 8 hrs before. (This was at 7pm). Surveyed in wet
conditions was not fun!

{[}Sketch survey showing Kamikaze and Lost Hopes{]}

\textbf{6-8-10 Jarv}

Jarv and Myles. After an epic 24hrs in camp waiting out the flood pulse
(hit @ \textasciitilde 7pm on 5-8-10), we are now getting ready to push
Korita. I wonder if others are coming from the surface\ldots{} We'll
have to rotate back on to the night train if a whole roster of 4 turn up
for the dream-team. Camp is super comfed with the roll mats from Cactus
Jnc installed under the tent.

\textbf{6/08/10 William}

James and William. Pushed a wet, so far unnamed pitch off Prince Consort
Road, could see what might have been the bottom about 10 m down at which
point the streamway really started to pour down (must be raining again).
Eventually bolted to the apparent bottom which of course wasn't, water
continued to hammer down a large hole in the floor, might have been a
challenge to rig and stay dry, but fortunately found a dry rift that
went over the active pitch (possibly). Another 40 m down that should be
easy to descend.

\textbf{6-8-10 Jarv}

Jarv and Myles. Train crash! Having rotated onto the day-train we
collided with Tet, Fratnik, JKP and William. So we must away, a long way
to the bivi, at least, when you leave at two minutes to midnight\ldots{}
Korita is dead, without explosive. Derigged entirely to the
3\textsuperscript{rd} pitch.

{[}Jarv's sketch topo of bottom of Korita{]}

\textbf{6-8-10 23:48 Tetley}

Back again for my seventh night! Fratnik and I had a smooth 2hr journey
down. Met Nicholas and Thara just about to head out. Jarv and Myles are
here too (soon to go down Korita). James and William then turned up. TEA
all round! Fratnik and I eventually went for a tourist jolly to the
Queen's Bed Chamber. then we went down to push Povodni Mož. Wow! A small
hole in a large dry passage soon leads to small fossil stream passage
then to nice Active stream. Went down 2 pitches (pushed by Izi, Erik,
Tjaša and Mawr). Bagged one more then half bolted another. Back to camp
for tea (and medals). It's clearly raining outside -- \emph{Zimmer} is
very wet. Hope Jarv and Myles (now back from pushing Korita) make it out
OK!

\textbf{7-8-10 9:50 a.m.} \textbf{Tetley}

Tea cooling down\ldots{} Fratnik is making food, James is putting his
wetish furry on, Dirty Old Town playing on sterio.

\textbf{7-8-10 11:10} \textbf{a.m. Tetley}

Final hot vitaminski -- then James and William to \emph{Serpentine},
Tetley and Fratnik to Povodni Mož. `Dangerous Dick' is seeping slowly
into my subconscious\ldots{}..

\textbf{7-8-10 11:15 p.m. Tetley}

Fratnik and I are back and well fed. A great days caving. Sadly Povodni
Mož is now dead, Up stream pushed to 30 m+ aven, bolting downstream, we
eventually hit a sump. All now surveyed and derigged. Now time for bed.

\textbf{8.8.10 Fratnik}

Drugo leto spet, Do konca nikoli izhojenih poti.

\textbf{8-8-10 10:40 a.m. Tetley}

So\ldots{} 8 nights I've spent down here -- a what a great time I've
had! Hopefully next year will be as fun as this year! Thanks to all my
fellow inmates at Camp \emph{X-Ray} -- for the tea, the jokes, the
laughs etc. Tomorrow is the 9\textsuperscript{th} Birthday of Friendship
gallery! Where will we be in 9 years time???

\textbf{9-8-10} \textbf{19:15} \textbf{Jarv}

Jarv and Thara. Alright chaps and chapettes we're taking the two packed
sacks and an extra with broken/superfluous tacklesacks and have unpacked
rope and fettled the armoury in the entrance. Grabbed a bit of rubbish
and Tet's Nido tin. I imagine Tuesday wave will be with you around Noon
Tet, James KP, William and Nicola are in it.

p.s. Recharging your MP3 player -- so long and thanks for all the fish!

\textbf{9/8/10 21:40 Kate}

Lie in till 4, Blackadder GOOD, throwing up BAD. Just come back from
investigating The palace of King Minos, the Queen's Bed Chamber and such
places. Did intend to push but it turned out that Dan had made up his
lead and his descender is absolutely obliterated so no rigging was done.
Had a lovely trip, nice climbs and awesome formations. Mud whirlpools
and more crystals than you can shake a stick at = FAB. \emph{Cheetah} is
horrifically terrifying. Back in camp, hurrah! last night this year,
it's been good. Can't wait to come back next year if you all will have
me.

\textbf{10/8/2010 1103 Dan}

Well well well, the final UG log entry. Almost finished packing up camp.
Kate just set of up. Left down are. Many tins of fish. tea bags and
sugar. Gas stove and 3 bottles of gas (2 big and full, 1 small and half
full) + lighters. In a D Drum w carbide for dryness. A few soups
chocolates + nuts. Hmm.. I guess I should get changed and leave. So long
till next year: \emph{X-Ray} 3 -- revenge of \emph{X-Ray}.

p.s. I hope the Loperamide kicks in soon.

Further things left: Loads of rope, a few slings, some booze, 2 survey
tapes, a lump hammer, coal chisel.

{[}20100731-23-54-22-Jarvist Frost-Canon G5-CRW\_0405-Camp
\emph{X-Ray}{]}

{[}20100729-13-06-55 - Iztok Mozir - P7294686 - Camp
\emph{X-Ray}{]}