

\section{\texorpdfstring{\emph{Zimmer}
streamway}{Zimmer streamway}}

James and Dave went for a quick bounce trip to look at the undescended
pitch below \emph{Zimmer} chamber, which started with a traverse in a
rift part-filled with boulders.

Bolts for the traverse led to a point where a drop into a water could be
seen. On bolting and descending this, the landing was in a shallow pool,
with water coming from some way back in the rift, having been collected
on the floor of \emph{Zimmer} chamber after running through the boulders
from both waterfalls.

The water then ran from the pool to cascade down a series of short drops
in a clean-washed rift with a friable slabby roof/wall overhanging it,
and eventually a point was reached with a drop too long to free-climb
down, with the passage beyond bending to the left and seeming to carry
on descending.

This certainly seemed like it had some potential, and was interesting in
that (with the exception of \emph{Republika}) it was the first
decent-flow active streamway encountered at depth as part of a main
route, rather than a streamway crosing the passage appearing from and
going to nowhere accessible.

A pleasant little bounce trip with an interesting and attractive find.

James, Jana and Dave returned a couple of days later to push a little
and take photographs, and while James and Jana got to work, Dave went to
drop a rope down to \emph{Falls Road} from \emph{Friendship Gallery},
since the Capt. K. crew had seemed likely to have reached there at a
lower level a day or two before, and a rope would enable a through-trip.
On descending, their anchors were found, showing that they actually had
been where we had calculated.

\name{Dave Wilson}
