\subsection{26/7/08 - Vilinksa Jama (AKA Tetley's mysterious
lead).}

Tetley, Dan, Bozo \& Janet (Guard).

Tetley led the way down past GW. Only grudgingly agreeing that blind
folds would not be required. The entrance is an unwelcoming yet
strangely inviting scree slope under a bedding plane, with a howling and
glacial draft emerging from it.

We headed down, leaving Janet at the surface to see off any would-be
unsurpers of the joys within. The way on quickly opens up to a chamber
with a snow plug then regresses just as quickly to a tight, boulder
filled, rift. Tetley started chisseling away at a constriction. Dan took
over while Tetley dug underneath, bypassing the squeeze entriely. More
up \& down around the boulders led to an impassably tight pitch head.
Tetley \& Dan made slow progress with chisels until Bozo arrived, having
extracted our location from Janet. A few mighty blows later, the way on
was clear (still tight though - will need more destruction).

Tetley rigged a ladder from a securely wedged boulder, and Dan clambered
down into a boulder strewn chamber. Water was dripping from a crack in
the ceiling, and the rift continued ahead. ``Does it go?'' shouted
Tetley ``Game on!'' came the reply.

Bozo forged ahead, quickly finding another chamber, filled with a vast
two way boulder slope. Up and over choked quickly. Down and under was
precarious in the extreme, but Tetley's careful proress under the
hanging death and down a short climb yielded the next pitch. A
\textasciitilde 20 m drop into a large fault controlled chamber. With no
SRT kit or rope we made our exit, pausing briefly to discuss potential
names. We settled on Vilinksa Jama (Veela cave) and returned to the Bivi
for Tea \& Medals. \name{DanG}

\fullwidthbox{26.07.08  Andy and Jana $\rightarrow$ M16 $\rightarrow$ \emph{Plopzilla} $\rightarrow$ end of \emph{NCB} (\emph{Zebra} passage)}{

10 hour trip. 
Re-rigged the way down to Gladiators. 
Also re-rigg a small traverse just after club mig.

Jana had an epic piss on the ledge in the middle of *Gladiators' Traverse*.

We check out \emph{Plopzilla}, but have not taken the rope out.

Checked the end of \emph{NCB}. 
In Zebra, 2 ways on:

1. A small climb down (\textasciitilde6 m) need to be smashed, remove boulders
2. Small, tight crawl continuing up \emph{Zebra}

Have taken the Gladiator's old rope out.
Amazing trip!

    
\name{Jana}
}