\section{9th Aug - Dangermouse Continued}

\margininbox{Dangermouse}{
     \begin{itemize}
    \item Jarvist Frost
    \item Martin McGowan
    \item Paul Hutton
    \item Tim Osborne
    \end{itemize}}{\explo}

Despite being on several expeditions I have managed to always find
somewhere else to go caving, but despite memories of Rik and Marcin's
exploration of \passage{Captain Kangaroo} I was lulled into going down to
\passage{Dangermouse}. On the way down CK it was noticeable that there is a
certain lack of quality rigging - reminding me of the early \textit{Ben's Crap Lead}
\sidenote{Former name of \passage{Vrtnarija} a.k.a. \passage{Gardeners' World}}
rigging. \passage{Bonus Chamber} could do with a traverse line before you
swing out over the pitch. The near two sections of the cave
\passage{Scrotty} and \passage{Even Scrottier} again with a few hours of bashing and
reshaping could change the passage to something nearer normal type
passage.

At \passage{Dangermouse} I headed down to Jarv's bolt and set about
clearing the pitch. Several large boulders were pushed down to a large
resounding boom and crack of shattering rocks. From the first bolt I
continued the gardening to a large natural in a corner. Paul followed on
down, and then Jarv. From the natural we descended via one bolt to a
window. A small moment of disco leg occured when a large chunk of the
wall came off as I stood on it.

Paul came down to the window and took over the bolting. Eventually,
after the first attempt broke the wall we had a bolt. I then sent /
volunteered Paul to traverse over the pitch head to a possible rift /
alcove. Unfortunately it turned out to be an alcove, so after a quick
piss Paul returned, and went down the pitch to the bottom. He declared
it dead \& I followed as far as a ledge above, then Jarv went to the
bottom for his compulsory photo op. Finally Tim joined us. While waiting
on Jarv to finish his photography I started to throw stones down a small
squeeze.

From beyond the squeeze a resounding boom could be heard. So quickly the
rope was pulled up leaving Jarv and Paul stranded below and Tim went
down. Unfortunately he faced the wrong way and could not turn his head
to see down the pitch. Putting my stop on my cowstails I descended for a
quick look. After the squeeze the cave opens out into a chamber.

So while Paul and Jarv surveyed out Tim placed his first bolt and rigged
the descent. A first for CK we backed up the descent a figure of eight
to a bolt and then an alpine butterfly onto a sling round a natural. All
of this beautiful rigging was executed by Tim. So we made quick visit to
the continuation and can confirm it is going somewhere.

Further Thoughts: We need a demolition team to visit \passage{Scrotty} and hammer,
beat it into shape.

\name{Martin McGowan}
