\section{\texorpdfstring{\emph{Plopzilla} - Photo Trip Writeup -
Jarv}{Plopzilla - Photo Trip Writeup - Jarv}}

Andy's last trip; so it had to be \emph{Plopzilla}.

Set off with a Daren drum of gear, with Paul in the lead and Andy
bringing up the rear (Ooh-err!) Took a few firefly digital shots coming
up into \emph{Hotline}. Seems to be lots off the lower \emph{M16}
entrance series, such as pitches into \emph{Lost City}, etc.

More photos on the newly sexy gladiator's rigging, then shot off up
\emph{Fawlty Towers}. Popped into \emph{NCB}; Andy conned us into
pissing down \emph{Silos}:- ``Don't piss in the main passage, climb down
there but no further than the first bolt!''

Rerigged traverse in \emph{NCB} (more photos!), then to the main game:
\emph{Plopzilla}. We had a plan, I asbeiled down with gear \& scrambled
up a slope. Pitch head was `interesting' (clip in cows, push through
squeeze and swing out above the 105 m drop hoping cows were still in).
Rebelays were `interesting' {[}so tight, one had to climb above belay to
derig descender{]}. Knot change `exciting' {[}knot a `Rik Special', 9 mm
below super slippery{]} Gear down there `intriguing'
\sidenote{this explains where all the rope went, 90 m 11 mm, 90 m 9 mm and 1 tacklesac}

We tried our best on 2 attempts at filming exploration, a few digital
pictures, and then dived into the Boulder choke. Went down -30 m or so,
drips, went through a lot of very loose Boulders then into bigger
stones, but with solid wall, lots of Boulders. Many drips, no stream
hearable. Otherwise very similar to \emph{Jelly Chamber}. Got as far as
a loose committing climb into a large-ish chamber.

It goes, it really does. But I'm not going there. Stones rattled
\textasciitilde 4 seconds.

Climbed ridge to corner and found another entry to choke.

So we left \& derigged. Rope → Club Mig. Sped out once, gear stashed.
2hrs from bottom to \emph{Fawlty Towers}.

\name{Jarvist Frost}
