
\section{Prospect for 2013}

Despite the success of the 2012 expedition, it is clear that there
remains a large amount of cave to be discovered in \emph{Vrtnarija} and
Migovec. In \emph{Vrtnarija}, the obvious areas to focus our attention
on include: the descent of the wet pitches at the end of \emph{Brezno
Slapov} and \emph{Invictus}, pushing the rift at the end of \emph{Milky
Way}, further exploration of \emph{Xanadu} and \emph{Lower Pleasures},
various unpushed leads in \emph{Watership Down} at the bottom, and a
second bolt climb to a window at \emph{Queen's Bedchamber} that seems to
be the continuation of the horizontal development found in Milk Way.
Furthermore, as we now have the necessary know-how and equipment to bolt
climb, there are various ¡deep¢ leads to revisit from the 2003
expedition such as Strap on the Nitro.

In some ways, we have the fortunate problem of almost being too
successful in recent years, as the leads are now fairly spread out and
getting further from camp each year. This year, it was common for
pushing trips to \emph{Watership Down} and
\emph{Invictus}/\emph{Atlantis} to take upwards of 21 hours.
Nonetheless, Camp \emph{X-Ray} remains the most likely camp due to its
central location and easy access to water -- however, an additional
lightweight 2 man camp nearer to the more remote locations is a strong
possibility.

Thanks to the connection there are also old leads in the old SysMig that
are worth revisiting, especially around the \emph{Waterloo} area.

In October 2012, on the annual ICCC/JSPDT ¡winter action¢ weekend, a
breakthrough was made in \emph{Primadona}, another cave system on
mountain. Due to its close proximity to the System, this will be the
next big connection project in our bid to shed more light on the complex
subterranean system under Migovec.

Acknowledgements

The authors are grateful for conversations with the entire Imperial
College and JSPDT expedition teams.

The Sledi Vetra 2012 expedition was funded with assistance from the Ghar
Parau Foundation and Imperial College Union Tours Fund (Recreational
Clubs Committee). We are indebted to the PD Tolmin and the Triglavski
Narodni Park (TNP) for continued support and permission to organise the
summer cave exploration camp on Tolminski Migovec.
