    \fullwidthbox{12-8-08 - \passage{Olympic Rift} - Jarv \& Tim}{From olympicrift.1, squeeze to pitch is SE, 5 m+. Wet sounding pitch
(Not Concorde?)

{[} Diagram of beyond \passage{Olympic Rift} .1 {]}

From last pitch, SW to aven (draft) + possibly over the top to pitch.
From scrotty.37, W down tight rift to ???

Squeeze is: a) Tight b) Bedrock either side c) no room for chisel. Bang ideally.

\name{Jarvist Frost}}




\section{Scrotty Bashing}

\margininbox{Scrotty}{
     \begin{itemize}
    \item Paul Hutton
    \item Martin McGowan
    \item Tim Osborne
    \end{itemize}}{\explo}

\recipecorner{Slop Kaboom}{
    \begin{enumerate}
    \def\labelenumi{\arabic{enumi}.}
    \item Into a pressure cooker put some tuna, soya mince, chilli paste, tomato puree, dried onion and water
    \item Leave cooking and go to sunset spot, have a few glasses of wine
    \item Let off some of the steam and take the pressure cooker off the heat
    \item Slowly try to turn the handles - if it won't go depressurise more
    \item Get fed up with this depressurising process and then force open the cooker
    \end{enumerate}    
    --- \textbf{KABOOM} ---
    \begin{enumerate}
    \def\labelenumi{\arabic{enumi}.}
    \setcounter{enumi}{5}
    \item Slop is instantly served - dinner can be got from all the surfaces - tarps, rocks and the chef
    \end{enumerate}
    Safety Advice:
    \begin{itemize}
    \item Listen to Janet's comment from last year log book
    \item Lost the weight from the cooker
    \item Do NOT try this at home/bivvy
    \item Wash down the bivvy afterwards
    \end{itemize}
\mininame{Martin McGowan}}


After the last trip, it was quite clear that the passages in
\passage{Captain Kangaroo} needed some alterations. So over a few drinks
and the fire I talked about the possible home improvements we could
make. The idea of bashing the shite out of \passage{Scrotty} seemed to appeal to
lots of people, no doubt motivated by the idea of getting something back
on the annoying part of the cave.

So Paul, Tim \& I set out armed with two hammers and chisels, as well as
some tatty rope for the climbs in the cave. We were quickly down at the
start of \passage{Captain Kangaroo} and we then set about bashing the
passage. We took a cycling approach to the task, with one person taking
the head of the peloton with the large hammer, then the second person
refining the passage with the smaller hammer and finally the third
person `resting' at the rear.

At \passage{Bonus Chamber}, Paul \& I rigged a traverse line, so you no longer
have to swing off the pitch whilst unprotected. On the same vein we made
several improvements to the handlines. The only one pitch that we could
not improve was the one that the the backup was below the main bolt. I
had removed the backup with the idea that on the way out we could rebolt
it with the kit left in \passage{Olympic Rift}.

Anyway after redecorating \passage{Scrotty} we listened to Bob and then headed
down for the kit left by Rik in 2007. We quickly descended the pitches
below \passage{Traverse Chamber}. It was quite clear from the offset that
this was not the normal Rik Venn rigging. It was a delight to go down
pitches with traverse lines, backups and other aspects of which would
pass for standard rigging from Alpine Caving Techniques.\sidenote{This sturdy hardcover manual, translated into English from the original French in the early 2000s, is a classic tome detailing vertical caving and rigging practices. No caving club is complete without a copy.} Well down we
went and it soon became clear that we were still in the \passage{Scrotty} series
according to PSS left by Clewin. At the bottom there was a rift going
off - was this the \passage{Olympic Rift}? After a few twists and turns it
was quite clear that this rift was very tight and we were fed up with
this shit. Paul had a quick look around the corner to check that the
rift continued in the same shit manner and then we got the feck out of
there.

Overall there is a new improved \passage{Scrotty}, but we have no idea where Rik's
lead is from last year. Maybe looking at the survey will solve this
problem.

\name{Martin McGowan}
