\subsection{Deep Push to Republika}

A lightweight bounce-push was made to -734 m to look at a major lead
left at the end of the 2004 deep camping expedition. Due to a
miscommunication during 2004, the upstream `tube tunnel' sized passage
was left unexplored until the last camping trip, everyone having assumed
that `Cow' died with a sump. This was pushed upstream over a few small
cascades to reach a large watershed (\emph{Republika Palma de Coco})
with the water entering from a high aven, literally splitting on a
triangular rock. It is hypothesised that this water could be coming from
\emph{Tolminska Korita}, 130 m away in both vertical and horizontal
directions. From the watershed, the new passage was followed downstream
to a 13 m pitch which was rigged with the tiny amount of gear that the
party had with them. This took them to a large active rift development,
with an unplumbed depth estimated at being in the region of 30 m. With
no more rigging gear, they surveyed out. The plan of this area is
interesting, with the two streams describing a crescent, with the end of
the rift in the new area being just 16 m away from the previous sump,
indicating that this sump is almost certainly perched and may feed onto
the same pitch series.

This was an extremely happy discovery for this cave, as the maximum
depth of -802 m is believed to be some 200 m above the shale-band
controlled water-table, and an active pitch has the potential to punch
through the sandy choked passage that otherwise dominate this area of
horizontal development.
