\section{Raw Logbook Writeups}

Typed by Jarv, 23-26th Dec 2011.

\subsection{Day 4 - Tuesday - James KP (by
handwriting)}

Andy \& James KP: Set off at 1-ish to rerig \emph{Laurel}. Got to the
top of \emph{Laurel} \& found the ropes, but not maillons! Decided to
walk back to bivvy \& look for more. Having found none, we left for
\emph{M2} to look for Clew \& Dan \& their stash of maillons. After a
LONG search for \emph{M2}, we reach \emph{M2} \& start going thru loose
meanders, constantly worried that we were going to shower Dan \& Clew in
rock fall. Eventually we get the maillons, walked back to GW, finally
rig \emph{laurel}. The second rope on \emph{Laurel} is slightly worn \&
could do with changing (15-20 m rope needed).

\subsection{\texorpdfstring{\emph{Planika} (Team JJ) -
Jana}{Planika (Team JJ) - Jana}}

Rigged down the end of the Rift and 1st Pitch in \emph{Planika}.
Pleasant surprise in the first chamber. The snow slope was quite melted
next to the pitch wall, so we were able to go all the way down. We
noticed a way on and started smashing the ice wall, feeling the draught.
We manage to get through, but after few meter it choked again with the
ice coming very close to wall. Back on the snow slope. On top we noticed
that another meltwater pit leading to the bottom of the ice, but this
needed bolting to get down, so we decided to continue the way. But not
for long, because a lot of fresh snow blocked the way forward. After
digging a bit through it we decided that a different route would yield a
way on. We were very cold as well.

Back on the surface, there was still time for me to chop down and
prepare some wood for the fire. Jarv walked down to Ravne to carry some
more stuff up.


\subsection{\texorpdfstring{\emph{M2} - Dan \&
Clewin}{M2 - Dan \& Clewin}}

After the obligatory morning faff in the bivvy we started off for
\emph{M2}, then turned back after 20 metres to get spits and cones.
Wandered over to the entrance which is a stone's throw from \emph{M18}.
The lower entrance drops into a gorgeous daylight shaft with a snow plug
in the bottom and vast amounts of scree.

We then followed the obvious way on, which rapidly gets less and less
obvious, turning into awkward rift. There are a fair few bits of free
climbing before we spotted a piton. Not wanting put all our trust into a
30 year old rusting bit of metal, we put in a couple of dodgy bolts as
back-ups to the piton. It was a straight drop, so we just used a few
metres of rope instead of one of the ladders we had. At the bottom is
ever squeezy and awkward rift ending in a ladder drop (another new
bolt). While doing this, Andy \& James appeared, begging for maillons.
We had far too many anyway, so we sent them off happy. At the end of the
ladder drop awaited yet more rift and free climbs. The last climb we
rigged with ladder before taking it off again since it didn't seem to
buy us anything. There it opened out into a reasonable, i.e.~more than 2
m, pitch. We put in two more bolts and then ran into trouble on the
third. Dan bludgeoned the spit teeth into a pulp and we spent about 10
minutes trying ingenious ways to remove the duff spit from the driver.
But we didn't want to do anything which would make it too difficult to
remove if things went wrong. It was already 7:30 pm so we abandoned the
mission.

\name{Clewin Griffiths}

\subsection{Cpt Kangaroo: Jana, James (Huggett?),
Tetley}

Easy down to \emph{Laurel}. First rebelay under the boulder turned out
to be too tight for our shorter members: add a long sling.

Down to \emph{Pico} no problem, then to \emph{Bonus Chamber},
\emph{Scrotty} would do with some mechanical enlarging. Down to
\emph{traverse Chamber}. Left 20 m of rope, 5x spit, maillon, hangers,
cones. Rigged a free climb to mud slump (could be improved).

Jana: First trip for me down to \emph{Mud Slump}. Lovely trip. It is
actually not that scrotty - at least, not for me\ldots{} :) \emph{Pico}
is an amazing pitch. Will definitely return down there!

\subsection{24.7.07 Jana}

Jana, Jarv --\textgreater{} \emph{Planika}

Went down and bolt in the first snow chamber to go down the new way,
where the snow has melted. \textasciitilde 8 m down in the snow tube
hole. The way on was squizy and tight, but a lot of draught blew in our
faces. We ran out of rope and bolts, so the lead it is still going,
albeit a bit tight.

We surveyed the new cave on the way out: we had gained 24 m of depth. We
still needed to go back and check the original pushing front from 2017,
which was a cold snow chamber.

We had almost finished rigging when we heard: ``Is the rope free?'' It
was Tetley - he had finally come down to \emph{Planika}. He wanted to
head down and take a look at the new going lead. We mentioned to him
that it was tight, which he shrugged off with a ``yeah, yeah\ldots{}''.
He still got stuck in the first rift further down. Not managing to
wrestle free immediately, he lit up a cigarette and started smoking it.
:) On the way out we derigged the \emph{Planika} ropes, to provide the
necessary lengths to rig down to \emph{Monatip} entrance.

\name{Jana}


\subsection{25-07-08 - E1 - Jarv}

Jana \& Jarv returned to E1, to investigate the possibilities\ldots{}

Jana pushed the small crawl, which was draughting went for a while.

We then moved attention to the deep deep crawl, abandoned last year as
dead. It wasn't. But perhaps it should have been. We moved the most
dangerous boulder loitering over the edge and set to work.

Dig dig dig\ldots{}

6 m long, \textasciitilde 60° scree slope, terminating in pitch (6-8 m,
as judged by rattling stones).

First few armfuls of rock were stacked, until we started chucking scree
over the edge, boom boom boom Tidied back to bedrock --- too tight ---
damn! Strange pitchead, almost T-shaped: water slopping over the edge,
but also a notch cut down (\textasciitilde10cm wide).

E1: We surface surveyed and tied the centrline to \emph{M2} entrance.
The end of the cave is 40 m above the end of \emph{Goodybag}, \emph{M16}
We will need a better name before it gets too big → `Mountain Goat
Cave'? Sounds good in Slovene, says Jana.

\name{Jarvist Frost}


\subsection{\texorpdfstring{25-7-08 \emph{Planika} →
EchoRift}{25-7-08 Planika → EchoRift}}

Dropped new melted pitch. Safe rock → awful rope rub. Meh. 1 bolt \&
down, comedy sling backup \& pitch rope stretched across. Pushed to
tight chest-wide rift; ice everywhere. Tight rift meandered to
pitchhead. Located natural hang while Jana fetched survey kit \& rope. 1
dodgy thread \& down. Long sling to stretch out rope (\textasciitilde8
m). This proved troublesome on the way up. Dropped onto ice-floored
chamber with semi-tight rift off it. 1.5 m diam hole to small chamber 4
m below. Even tighter rift in same dirn. from here. Surveyed out
carefully.

Just the otherside of the chest squeeze, Tetley shouted down, having
taken the hyper death slide across the snow. He tried the chest squeeze,
got wedged \& so cracked out his smoke kit \& had a tab.


\subsection{\texorpdfstring{24/7 \emph{M2} Rigging Part 2 -
Clewin}{24/7 M2 Rigging Part 2 - Clewin}}

Dan \& Clewin: Started where we left off --- with 2 ½ bolts placed at
the top of the first real pitch. Dan finished off the last bolt and so
we abseiled down. Using the rope from the pitch I climbed up from the
bottom to see if there was a way on at a high level. There wasn't, but
the aven looks promising, although would require a proper climbing trip.
The bottom of the pitch leads to a 40 m pitch. There was a 70s piton for
backup and a spit at the pitchead. We used the spit (not \emph{too}
rusty) and put a new spit to give a Y-hang. 30 m down was a wide ledge
with garbage from the 70s.

{[}Side track{]} Off the ledge was a small rift leading to a drop which
we rigged with a ladder. Below that a dodgy free climb lead to a nice
section of wide rift passage with a stream at the bottom and a red
survey dot on the wall. This all closed up completely, so we derigged
the ladder.

At the bottom of the 40 m (Kletnikov Skropilnik) a tight rift of black
rock lead to the next pitch. There were no pitchead bolts so we left
this for another trip. Clewin.


\subsection{\texorpdfstring{26/7 Fratnik \& James: 3rd Rig of
\emph{M2}}{26/7 Fratnik \& James: 3rd Rig of M2}}

Continued where we left: Andrej rigged \emph{M2}: safely 1st rebelay on
a giant (minibus sized) boulder of dubious stability. 2hrs + 5 rebelays
later we were 20 m from the bottom \& out of spits. Damn! Return with 5+
bolts and finish the job some other day. Had a hell of a time getting
out of the entrance rift\ldots{}
