\section{Expedition Proceedings}

During the 2008 expedition a consider effort was spent exploring the
\emph{Captain Kangaroo} branch of \emph{Vrtnarija}, with the sole aim of
connection to the passage in Sistem Migovec. Our rate of exploration was
limited by the length of time taken to get to the pushing front, with
fifteen hour trips being the minimum to achieve much. The obvious
solution for 2009 was to setup a camp.

Thirty-two people-nights were spent at the two man underground camp,
with two twelve hour shifts of hot-bedding. Exploration in this section
(\emph{Captain Kangaroo}, \emph{Vrtnarija}) was split between looking at
climbing leads in the hope of connecting to \emph{M2}, and pushing the
\emph{Dark Tranquillity} series downwards. The majority of the climbing
was done via bounce trips from the surface, and also as `light days' for
people on their way out from camp. All the deep pushing was done from
the camp downwards. Total amount of caving was 98 people trips,
including camping trips.

In 2008 we had left the developing pitch series at the bottom of
\emph{Dark Tranquillity}, as this level in \emph{Vrtnarija} had now
dropped below the known extent of \emph{M2}; the main expedition aim in
2009 was to try and connect \emph{Vrtnarija} to Sistem Migovec. Cave
exploration progressed quickly and easily, with depth building at the
rate at which pitches could be safely rigged, gardened and surveyed.

Within two weeks of the caving commencing we had bottomed \emph{Happy
Monday} (P81 m), with the resulting survey data indicating an almost
inevitable intersection with the horizontal development at -550 m in
\emph{Vrtnarija} (\emph{Friendship Gallery}). The next team down pushed
through the terrifying boulder choke at the floor of the chamber
(\emph{Hanging Garden}), gained a well developed rift with phreatic
crawlspace at the top, and pushed downwards until they found an old bolt
just beyond a small confluence.

Plotting the data back on the mountain top it was clear that they had
intersected `\emph{Falls Road}', a set of tight active rifts accessed
just below \emph{Friendship Gallery}. The next trip into the cave went
via the main pitch series and dropped a rope down from this \emph{Prima
Junction}, intersecting the rigging left by the previous push. Strangely
the 2001/2003 explorers had missed the lead towards `\emph{Happy
Monday}', even though the end of the large rift forms one of the
original rebelay bolts for the confluence pitch.

The last camping trip saw an epic list of `must do this year' tasks,
dropping into the newly discovered \emph{Tolminska Korita} via the
easily passed main pitch series, taking pictures of new discoveries and
undocumented pitches along the way, recording the new discoveries from
\emph{Falls road} back to \emph{Walk the Line}, resting overnight at
camp, then descending to the connection point once more to finish off
the survey, investigate the `\emph{Muddy Window}' phreatic passage 10 m
above the floor of \emph{Happy Monday}, and derig back to camp before
resting again and then striking camp in the morning.
