\section{Imperial College Press Release}

This summer Imperial College Caving Club discovered the longest cave in Slovenia during their Sledi Vetra 2012 expedition.

The Imperial cavers have organised joint expeditions with the local Slovene club (JSPDT) to the mountain of Tolminski Migovec since 1994, with more than 80 Imperial students having contributed to the discovery. Every year they explored deeper and further into the mountain, the main discoveries being two large and deep cave systems (System Migovec and System Vrtnarija), both notable caves in their own right. A further 2000 m of cave passage was discovered this summer, leading to the connection of these two cave systems at a depth of 650 m.

The combined cave system is 24.9 km long and 975 m deep. The deepest point was found this summer where the cave passage leads down into an extensive crystal clear flooded section. The vast majority of such long cave systems in the world are shallow and warm, which greatly eases their exploration. The Migovec caves are deep, extremely vertical and cold, requiring a high degree of technical skill and physical endurance.

Tolminski Migovec is on the edge of a major thrust complex, the Slatna overthrust, with steeply dipping faults cutting and offsetting the thrusts. A complex series of faults and folds have led to extremely complicated cave formation in the Triassic limestone, the computer model of the mountain’s cave passage looking like a piece of shattered glass. Understanding the hydrology of the mountain is significant as it is the watershed of the Adriatic and Black Seas.

Deep exploration is made possible by the team remaining underground for 3-4 days at a time, sleeping over 600 metres below the surface of the mountain. As well as the physical challenge, they have had to innovate in developing new methods and techniques – for instance in climbing with a lightweight cordless drill adapted for caving and fitted with external batteries recharged by solar power.

The connection of these systems, and the discovery of the longest cave in Slovenia has been the major effort and activity of the Imperial cavers for the last five years. But the exploration of Migovec is certainly not over – every new passage reveals a new place within the mountain, and no one can possibly tell what will lie around the next corner.