copied from HM3, slowly being edited as of 07.11.2023





Acknowledgements 2012

The authors are grateful for conversations with the entire Imperial
College and JSPDT expedition teams.

The Sledi Vetra 2012 expedition was funded with assistance from the Ghar
Parau Foundation and Imperial College Union Tours Fund (Recreational
Clubs Committee). We are indebted to the PD Tolmin and the Triglavski
Narodni Park (TNP) for continued support and permission to organise the
summer cave exploration camp on Tolminski Migovec.



\chapter{Sponsorships and thanks}
We cannot thank enough the following organisations who have supported the Migovec expeditions over the years, and in particular from 2007 to 2012.
\begin{citemize}
\item The \textsc{Ghar Parau Foundation} for the \textsc{Ghar Parau Expedition Grant Award}, which provided funds for buying all kinds of equipment (especially rope!) -- and its subset, the \textsc{Alex Pitcher Award} which helped young cavers buy their caving helmets and lights and allowed them into expedition caving. This ongoing support played a pivotal role in promoting a thriving club community.
\item \textsc{Beast Products} for their sponsorship and donation of thermals for underground camp in 2009, which were put to use again and again in the following years and without which a good night's sleep underground would have been much less likely.
\item The \textsc{Imperial College Trust} and  \textsc{Imperial College Union} for Tours funding for the financial support directed at transportation subsidy.

\item The 2010 and 2012\textsc{Viljem Putick Prize} awarded for best Slovenian exploration project, put to good use in the purchase of additional rigging equipment.

\item \textsc{Felix}, the Imperial College weekly newspaper for the publication of exploration stories.


\mydelimiter

\item All the editors of this volume of the Hollow Mountain, more than a decade in the making - \textit{Jarvist Frost} for starting it all, \textit{Ben Richards} for getting it going again, and \textit{Tanguy Racine} for providing the template from the excellent Hollow Mountain 3, which this volume merely emulates. 

\item All photographers and authors who contributed to this project, and particularly to those who meticulously labelled their photographs - \textit{Jarvist Frost}, \textit{Jana Carga}, and \textit{Izi Mozir}. A special thanks goes to to \textit{Thara Supasiti} as the only cartoonist in this volume.

\item \textit{Tony Seddon} and \textsc{Starless River} for supplying expo equipment (often with a large discount), rope rescue practice sessions held in the UK, and general advice on everything expeditionary.

\item \textit{Mark} and \textit{Jim Evans}, for starting it all for ICCC, and for coming back often to share stories of early exploration at the sunset spot.
\item The \textit{lags}, for want of a better name, who return to Slovenia year after year to explore cave passage and share their knowledge.
\item \textit{James 'Tetley' Hooper}, for the use of his flat in Tolmin as a base throughout 2007--2012.
\item \textit{Janet Cotter}, for doing a million tasks in the Bivi and always cracking a smile for returning hungry cavers, and years of surface bashing work. 
\item \textit{Dave Wilson}, for the design and sale of many Bisun lights (nee Mig lights) to new cavers, for the technical expertise in setting up the solar panel electronics, on which the expedition depends.


\item \texit{Jarvist Frost} again for the colossal amount of work dedicated to the Hollow Mountain project, and every word and Git command that forms the basis of this book. 


\item The \textit{Klobu\v{c}ar} family: Slavica, Marjan, Nada, Zoran\ldots{} for the generous use of the barn to store equipment and park the van during the expeditions.

\item Our friends at the JSPDT, with whom we've shared the toil and the rewards under Migovec.
\item The \textsc{Jamarska Re\v{s}evalna Slu\v{z}ba} (JRS), the \textsc{Jamarska zveza Slovenije} (JZS) whithin which it operates, the helicopter pilot from mountain rescue services whom we must thank again.
\item \textit{Zdenko Rejec}, for heading the JSPDT and being so supportive. Hvala!
\item \textit{Jana \v{C}arga} for the work behind the scenes within the JSPDT, for translating Slovenian to English and vice versa over and over, and for every photograph.
\item \textit{Andrej Fratnik}, whose welcoming spirit sparked this joint venture and whose participation in the project from the very early start to this day has never faltered. 
\end{citemize}


Vrnili se bomo! (We'll be back!)