\section{Stores Sesh (8 weeks to go!)}

I texted Jan just after midday, he was still in bed. Chastising him
gently \& signing off with a usual “to minus 1000” I get by way of
reply:

“I’m visualising -2000 – mental preparation is key”

With a typicaly ICCC alpine start, the day dragged on into festernoon and the workers assembled.

We checked out all the club tents — we have two A.OK Mk 4 F10s, for
that classic orange vibe, and the Equinox 450 (The Casino) actually
seems OK (no broken poles). Though all could do with a proof.

Three sets of (11,10,10mm Mammut) 200m rope were carefully chained in
the quad, and put in the soak.

Finally, Jan spent the afternoon assembling bits of our Raumer
Stick-Up into a working setup for bolt climbing. Our plan is to check
it and the drills out in Yorkshire in a couple of weekends time. For
now, however, we’ve had to make do with a hilarious ‘climbing the
north face of Beit Quad’ with tent-pegs nicked from the tents as
pretend rawl bolts, checking that the lengths of the various etriers
agrees with what we can reach with the drill. 

\name{Jarvist Frost}


\section{T-7 days}

On the day the last ever space shuttle blasts off, we’ve entered that
horrible time seven days or so before expedition when it’s suddenly
become too late to sort out anything major, yet there’s lots of
trivial jobs that need doing… at some point. Just one last package
to arrive in the post, a new 80W solar panel to be used on the
mountain at the campsite, then transferred to the Slovenians for use
on the new mini mountain hut which is proposed to be flown up to Mig
this Autumn.

Still, last weekend we spent a full day around the ASDA and Lidl at
Clapham junction, before abusing a taxi to transport our swag back to
South Ken. After throwing away all the superfluous cardboard food
comes in, and packing it densely into green removals crates, we went
home exhausted. Returning on Sunday we cut rigging tape to length,
sorted all of our many (so many!) first aid kits and carefully doubled
bagged and densely stuffed underground camping gear into tackle sacks.
Our fairly sizeable stores are absolutely stuffed with crates, bags
and big blue barrels. How we fit it all in and on a 9-seater Transit I
will never truly understand.

Minor jobs proceeded during the week. This weekend we are finishing
off the shopping on Saturday (a hundred trivial things we can’t
survive without: peanut butter, soya mince, squeezy marmite etc.),
packing personal caving gear and sorting out all the rigging and
bolting equipment we need.

To this end my sitting room currently looks like some unholy cross
between caving stores and an electronics workshop. A smouldering
solder iron lurks in the corner, overseeing the contents of burst open
drills; the smell of rosin flux competing with the darker note of
Toluene rising off the patched oversuit on the floor. 

From an organisational point of view I find these times very strange,
everything slowly going quiet as the ToDo list shrinks, itineraries
are filled, all bags are packed. Underground camp bags are ready to be
opened at -550m, bolting kits are prepared, survey books are filled
with neat indelible guidelines, carefully washed and dried rope is
ready to spool out of bags while abseiling down hitherto undescended
shafts, shedding light on the walls for the first time in history.

Like some grotesque Hemulen in a Moomin novel just as the true raison d'etre of the expedition (the actual exploration of new cave) starts, the need for any further organisation disappears. Nothing left to do but lace up the walking boots, fill a bag and head up Migovec herself.

\name{Jarvist Frost}