\section{\texorpdfstring{Through trip \emph{M16} -
\emph{M2}}{Through trip M16 - M2}}

This year a few of us went up Mig; Đaljo, Špela, Erik, Tjaša and Karin.
As usually a stop at Kal is necessary. Once on top, it was already dark
and most of peoples were already asleep. The next day also Zec, Antonio
and Teja came up. We divided ourselves into three groups. Tetley did not
miss a chance to take Karin and Tjaša; who were freshers on Mig, caving
down \emph{M16} (\emph{Titanic}). On the end they ended up doing a
through trip and came out from \emph{M2}.

I took Teja to the top of \emph{Pico} and back. The others went to
\emph{M16}.

We were all out in time for a traditional sunset gathering, which is
truly unforgivable.

The next day Tetley suggested doing the same through trip
(M16-\emph{M2}) and at the same time check out one pitch in \emph{NCB}.
We liked the plan. The leaders were Karin and Tjaša followed by myself,
Erik and Thara. We all knew the way till \emph{Titanic}, but none of us
was in \emph{NCB} gallery before. We saw 3 pitches but were unsure as to
which pitch we need to check out. We went all the way till the end of
the gallery and on the way back decided to check out the middle pitch.
We bolted and Thara went down first to check it out. He came back up
quickly with the news that there is no way on. At this point we decided
best to carry on with our loop trip. We were at \emph{Silos} now and
Tjaša went first down. There is a big traverse connecting \emph{M16}
with \emph{Silos}. I somehow managed to get stuck on it. After 20 min of
struggling I managed to get free. After that \emph{M2} - there was quite
a lot of nasty meanders to go through. After 8 hours we were finally
out. Back in the bivy we were laughing which pitch we should check it
out\ldots{}

\name{Izi}
